Longitudinal Principal Manifold Estimation

### 2024-04-03 paper/main.tex
add aggregate results for larger adni population

Considering the same individual shown in Figures \ref{fig:adni_result} and \ref{fig:adni_cross_section}, results of fitting the PME and LPME algorithms to the surface of the left thalamus are shown in Figures \ref{fig:adni_lthal_result} and \ref{fig:adni_lthal_cross_section}. While the segmentation estimates of the thalamus do not show the same sort of deviations that are clearly visible for the hippocampus, the cross section plots shown in Figure \ref{fig:adni_lthal_cross_section} indicate that both the PME and LPME algorithms show a closer fit to the more regularly shaped thalamus surface than was achieved for the hippocampus. This is reflected in the smaller discrepancies between volume estimates, as seen in Figure \ref{fig:lthal_volumes}, though visible variations between time points remain.

@@

\begin{figure}[h]
  \centering
  \subfloat[\centering Left Hippocampus Cross Section]{\label{fig:lhipp_cross_sections} \includegraphics[width=8.5cm]{adni_plots/adni_cross_section}}%
  \hfill
  \subfloat[\centering Estimated Volumes]{\label{fig:lhipp_volumes} \includegraphics[width=8.5cm]{adni_plots/lhipp_volume}}
  \caption{Left Hippocampus Cross Section and Volume Estimates. In (a), we can see that the LPME-approximated embedding function, shown in blue, routinely falls inside the boundaries of surface in the observed data. In the fifth observation with $t=3.04$, the LPME-estimated surface has a gap in it, corresponding to an unexpectedly high volume estimate seen in (b). Volume estimates derived from the observed data are consistently higher than those obtained from the embedding functions of PME and LPME estimates, respectively.}
  \label{fig:adni_cross_section}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \subfloat[\centering Left Thalamus Cross Section]{\label{fig:lthal_cross_sections} \includegraphics[width=8.5cm]{adni_plots/adni_lthal_cross_section}}%
  \hfill
  \subfloat[\centering Estimated Volumes]{\label{fig:lthal_volumes} \includegraphics[width=8.5cm]{adni_plots/lthal_volume}}
  \caption{Left Thalamus Cross Section and Volume Estimates. In (a), the cross sectional plots indicate that the PME and LPME algorithms fit much more closely to the regularly shaped thalamus surface. The volume estimates in (b) show similar general trends to the hippocampus estimates, with the estimates reached using the observed data being higher than the PME- and LPME-based estimates. However, the differences between the estimates taken from the PME and LPME results are more comparable.}
  \label{fig:adni_lthal_cross_section}
\end{figure}

>>>>

Considering the same individual shown in Figures \ref{fig:adni_result} and \ref{fig:adni_cross_section}, results of fitting the PME and LPME algorithms to the surface of the left thalamus are shown in Figures \ref{fig:adni_lthal_result} and \ref{fig:lthal_cross_sections}. While the segmentation estimates of the thalamus do not show the same sort of deviations that are clearly visible for the hippocampus, the cross section plots shown in Figure \ref{fig:lthal_cross_sections} indicate that both the PME and LPME algorithms show a closer fit to the more regularly shaped thalamus surface than was achieved for the hippocampus. This is reflected in the smaller discrepancies between volume estimates, as seen in Figure \ref{fig:lthal_volumes}, though visible variations between time points remain.

@@

% \begin{figure}[h]
%   \centering
%   \subfloat[\centering Left Hippocampus Cross Section]{\label{fig:lhipp_cross_sections} \includegraphics[width=8.5cm]{adni_plots/adni_cross_section}}%
%   \hfill
%   \subfloat[\centering Estimated Volumes]{\label{fig:lhipp_volumes} \includegraphics[width=8.5cm]{adni_plots/lhipp_volume}}
%   \caption{Left Hippocampus Cross Section and Volume Estimates. In (a), we can see that the LPME-approximated embedding function, shown in blue, routinely falls inside the boundaries of surface in the observed data. In the fifth observation with $t=3.04$, the LPME-estimated surface has a gap in it, corresponding to an unexpectedly high volume estimate seen in (b). Volume estimates derived from the observed data are consistently higher than those obtained from the embedding functions of PME and LPME estimates, respectively.}
%   \label{fig:adni_cross_section}
% \end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_plots/adni_cross_section}
  \caption{Left Hippocampus Cross Section. We can see that the LPME-approximated embedding function, shown in blue, routinely falls inside the boundaries of the surface in the observed data. In the fifth observation with $t=3.04$, the LPME-estimated surface has a gap in it. Gaps in the surface estimates frequently correspond with inaccurate volume estimates.}
  \label{fig:lhipp_cross_sections}
\end{figure}

@@

% \begin{figure}[h]
%   \centering
%   \subfloat[\centering Left Thalamus Cross Section]{\label{fig:lthal_cross_sections} \includegraphics[width=8.5cm]{adni_plots/adni_lthal_cross_section}}%
%   \hfill
%   \subfloat[\centering Estimated Volumes]{\label{fig:lthal_volumes} \includegraphics[width=8.5cm]{adni_plots/lthal_volume}}
%   \caption{Left Thalamus Cross Section and Volume Estimates. In (a), the cross sectional plots indicate that the PME and LPME algorithms fit much more closely to the regularly shaped thalamus surface. The volume estimates in (b) show similar general trends to the hippocampus estimates, with the estimates reached using the observed data being higher than the PME- and LPME-based estimates. However, the differences between the estimates taken from the PME and LPME results are more comparable.}
%   \label{fig:adni_lthal_cross_section}
% \end{figure}
%

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_plots/adni_lthal_cross_section}
  \caption{Left Thalamus Cross Section. The cross sectional plots indicate that the PME and LPME algorithms fit much more closely to the regularly shaped thalamus surface.}
  \label{fig:lthal_cross_sections}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_plots/adni_lhipp_volume_comp}
  \caption{Left Hippocampus Volume Estimates for Six ADNI Participants. All volumes are estimated by summing the volumes of all voxels contained within the structure estimate as described by segmented data (red), LPME (blue), and PME (green). When PME and LPME are applied to the irregular shape of the hippocampus, there is a clear ordering of volume estimates, gaps in the estimated surface induce underestimates of the volume compared to the volume values estimated from the data. These gaps may also cause inconsistency in the LPME volume estimates, resulting in higher aggregate variability than for the data or PME estimates.}
  \label{fig:lhipp_volume_comparison}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_plots/adni_lthal_volume_comp}
  \caption{Left Thalamus Volume Estimates for Six ADNI Participants. Volume estimates are obtained by summing the volumes of voxels contained within the structure as described by segmented data (red), LPME (blue), and PME (green). The PME-based volume estimates closely track the volumes estimated from the segmented data, reflecting the close fit of the PME algorithm to the regularly-shaped thalamus data. The LPME estimates appear to successfully smooth over regions with large variations in subsequent volume measures from the data and PME estimates, as seen for participants 021\_S\_0626 and 053\_S\_0919.}
  \label{fig:lthal_volume_comparison}
\end{figure}

\begin{table}[ht]
  \centering
  \begin{tabular}{|c c c c|}
    \hline
    Structure & Data & LPME & PME  \\
    \hline
    Left Hippocampus & 218 & 225 & 191 \\
    Right Hippocampus & 221 & 245 & 195 \\
    Left Thalamus & 82.9 & 49.7 & 72.0 \\
    Right Thalamus & 147 & 116 & 130 \\
    \hline
  \end{tabular}
  \caption{Mean of the empirical standard deviation of volume estimates computed for each ADNI participant.}
  \label{table:adni_volume_sds}
\end{table}
